---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
library(dplyr)
library(bwplot)
library(ggplot2)
```

# bwplot

<!-- badges: start -->
[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)

[![License: MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
<!-- badges: end -->

The goal of bwplot is to generate plots with an aesthetically cleaner theme than the `ggplot` default. It is handy for quickly producing numerous plots that share the same general template and vary only in plot variables. By presetting certain theme elements in the `bwplot` functions, the need to repeat code for each individual plot is reduced.

`bwplot` uses the `theme_bw()` theme by default and removes gridlines and legends whenever appropriate. A generic colour scheme is also created. Axis labels, colour palettes, etc. can be edited using additional `ggplot2` functions after initially generating the plot.

This package is currently under construction as part of UBC's STAT 545B course for the Winter 2020 semester.


## Installation

You can install the development version from [GitHub](https://github.com/) with:

```r
devtools::install_github("t-wang-ecohydro/bwplot")
```
## Example

Here is a basic example comparing the a `bwplot` function to the equivalent `ggplot2` alternative:

```{r example}
# ggplot2:
mtcars %>%
  group_by(cyl) %>%
  summarise(mean = mean(mpg, na.rm=TRUE)) %>%
  ggplot() + geom_col(aes(as.factor(cyl), mean, fill = as.factor(cyl))) +
  xlab("cyl") + ylab("mpg") +
  theme_bw() + theme(panel.grid.major = element_blank(),
                     panel.grid.minor = element_blank(),
                     legend.position = "none") +
  ggtitle("Mean mpg per cyl")
```


```{r example2}
# bwplot:
bw_mean_col(mtcars, mpg, cyl, na.rm=TRUE)
```


## Available functions

Current plotting functions are:
1. `bw_mean_col`: produces a `geom_col()` plot with the following defaults:
  - Background: `theme_bw()`
  - Gridlines: none
  - Legend: none
  - Fill: unique color per column
  - Axis labels: default data table column names
2. (to be completed in later assignments?)


## How it was made (this part is for Assignment 2)

### Creating the package

The first step is to create the package. This was done using`create_package()`, as shown in the screenshot below:

[I took this right before restarting my R session](man/figures/A2_createpackage.png)

`bwplot` is the name of my package. An `R` folder for functions, `NAMESPACE` file and `DESCRIPTION` file were automatically generated. 

Next, I set up version control using `use_git()`, and made my first commit. On GitHub, I created a repository called "bwplot" and synced my new R package to it via the Git terminal. 

```git
$ git remote add origin https://github.com/t-wang-ecohydro/bwplot
```

Having finished the initial setup, I added my function from Assignment 1 to the package.

### Creating a function

I generated the .R file using:

```r
use_r("bw_mean_col")
```
This created `bw_mean_col.R` and the also prompted me to create a matching test file. (I got tired of taking screenshots by this point, sorry!) I took my function from Assignment 1 and pasted it into the new file.

### ROxygen comments

In RStudio, I navigated to my function and used "Insert ROxygen Skeleton" to generate a ROxygen comment section. Then I filled in the title, parameters, and additional information about my function. I used `@export` to indicate that this function is meant to be public (it's the main feature of my package after all), and added package dependencies using `@import` (see the next section). For this part of my package creation, I edited my R file manually.

I called `document()` to save my changes. Now when I use `?bw_mean_col` I can see a basic description of what my function does.

### License

I generated a `LICENSE` file and `LICENSE.md` using `use_mit_license(copyright_holder = "Tracy Wang")`. I went with the MIT license because it seemed like a good general license... this is just a toy package after all, I'm not particularly concerned about people grabbing it to use or modify.


### Package dependencies

My function has some dependencies on other R packages. To add these dependencies, I typed into the console:

```r
use_package("dplyr")
use_package("ggplot2")
use_package("palmerpenguins")
```

This updated the `DESCRIPTION` file to include these dependencies. When I used `document()`, the `NAMESPACE` file was automatically updated as well. In the ROxygen comment section of my function, I added my package dependencies:

```r
#' @import dplyr
#' @import ggplot2
```

For the `%>%` piping function, I tried the following:

```r
use_pipe()
```
But when I ran `check()`, I got an error stating that the dependency on the `magrittr` package was not needed, so I ended up doing it the other way by using `use_package("magrittr")` and then editing the ROxygen comment.

```r
#' @importFrom magrittr "%>%"
```

Finally, I went through my function and changed all functions with package dependencies into "package::function()" format. For example, `ggsave()` became `ggplot2::ggsave()`, and `select()` became `dplyr::select()`.


### Updating the description

I added my name and email manually under the "Author" field and changed the license. For the title, I experimented with editing the file via the console:

```r
use_description(fields = list(Title = "Make Aesthetically Tidy Plots"))
```

### Tests

I generated a `testthat` folder and a `testthat.R` file using the command:

```r
use_testthat()
```

Then I generated `test-bw_mean_col.R` using `use_test("bw_mean_col")`. I pasted my tests from Assignment 1 into the file. Because I had assigned some variables last time in order to test my function, I decided to change some of the tests to check for plot layer parameters, such as `labels`, `mapping` and `data`, rather than directly comparing whole plot objects to each other. 

Because some of my tests use the `penguins` dataset from `palmerpenguins`, I had added it as a package dependency (but I didn't need to `@import` it explicitly in order to use it, just refer to the dataset as `palmerpenguins::penguins`).


### Checks

Throughout the package creation process, I was constantly running `check()` to make sure that everything was working! At the very bottom section of this readme is what my most recent check looks like (it's a bit of an eyeful so I'm pushing it aside to move on to other sections).

### Readme and vignette

The last step of my assignment was to create a readme file and vignette via the console:

```r
use_readme_rmd()
use_vignette("how-to-use")
```
This generated a `README.Rmd` file (the readme) and a `how-to-use.Rmd` file (the vignette). I edited the content of both these files manually on RStudio.

### Optionals:

Code of conduct: I generated the file using `use_code_of_conduct()`. 

News: 

Package website: I used the following commands in the console:

```r
use_pkgdown()
pkgdown::build_site()
```
And lo and behold! My package now has a website that can be viewed [here](file:///C:/Users/Tracy/Documents/Uni/UBC%202020-2021/STAT%20545B/bwplot/docs/index.html).

### Now here comes that `check()` output

```{r}
# library(devtools)
# check()
```
